---
title: "Installation"
editor: visual
format:
  html:
    toc: false
---

## Installing R and RStudio

To work with the `goFlux` package, you'll need to install R and RStudio. Both programs are free and easy to download. For an easy step by step, follow the procedure described [here](https://rstudio-education.github.io/hopr/starting.html#using-r).

## Installing goFlux from GitHub {#sec-installgoflux}

To install a package from GitHub, one can use the package `devtools` (or `remotes`) from the CRAN:

```{r}
#| eval: false
if (!require("devtools", quietly = TRUE)) install.packages("devtools")
```

Then, install the `goFlux` package from GitHub. If it is not the first time you install the package, it must first be detached before being updated.

```{r}
#| eval: false
try(detach("package:goFlux", unload = TRUE), silent = TRUE)
devtools::install_github("Qepanna/goFlux")
```

::: callout-important
The package is constantly being updated with new functions or de-bugging. To make sure that you are using the latest version, re-install the package every time you use it.
:::

::: callout-warning
**If prompted, it is recommended to update any pre-installed packages.** The functioning of the package depends on many other packages (`AICcmodavg`, `data.table`, `dplyr`, `ggnewscale`, `ggplot2`, `ggstar`, `graphics`, `grDevices`, `grid`, `gridExtra`, `jsonlite`, `lifecycle`, `lubridate`, `minpack.lm`, `msm`, `pbapply`, `plyr`, `purrr`, `rjson`, `rlang`, `rlist`, `SimDesign`, `stats`, `stringr`, `tibble`, `tidyr`, `toOrdinal`, `utils`), which will be installed when installing `goFlux`.
:::

Once installed, load the package `goFlux`:

```{r}
#| eval: false
library(goFlux)
```

## Verify installation {#sec-verify}

You can verify that `goFlux` is properly installed and all dependencies are available:

```{r}
#| eval: false
# Check if goFlux is installed and can be loaded
if (requireNamespace("goFlux", quietly = TRUE)) {
  cat("✓ goFlux is installed and available\n")
  cat("✓ Version:", as.character(packageVersion("goFlux")), "\n")
} else {
  cat("✗ goFlux is not installed or cannot be loaded\n")
}

# Test that core functions are available
core_functions <- c("goFlux", "best.flux", "import2RData", "autoID")
for (func in core_functions) {
  if (exists(func, mode = "function", where = asNamespace("goFlux"))) {
    cat("✓", func, "function available\n")
  } else {
    cat("✗", func, "function not found\n")
  }
}
```

## Other required packages {#sec-packages}

Other packages from the CRAN are required to run the examples on this website.

```{r}
#| eval: false
# Install packages
install.packages("dplyr")
install.packages("purrr")
install.packages("readxl")
install.packages("openxlsx")
```

```{r}
#| eval: false
# Load packages
library(dplyr)
library(purrr)
library(readxl)
library(openxlsx)
```
