---
title: "Other functions in the package goFlux"
format: 
  html: 
    css: styles.css
engine: knitr
editor: visual
---

Aside from the typical workflow of flux calculation, one may wish to estimate other parameters from gas measurements. Currently, this page only contains few additional functions. To request more complementary functions, [open an issue on GitHub](https://github.com/Qepanna/goFlux/issues) or contact the maintainer of this package, Karelle Rheault ([karh\@ign.ku.dk](mailto:karh@ign.ku.dk){.email}), directly.

### `iso.comp` for isotopic composition determination

Extracts the y-intercept of the regression between the isotope ratios (13C, 15N or 18O) plotted againts the inverse of the gas concentration (CO~2~ or N~2~O).

#### Usage

::: callout-note
Code chunks under **Usage** sections are not part of the demonstration. They are meant to show you how to use the arguments in the function.
:::

```{r}
#| eval: false
#| code-copy: false
iso.comp(
  dataframe,
  CO2dry_ppm = "CO2dry_ppm",
  delta_13C = "Delta_Raw_iCO2",
  N2Odry_ppb = "N2Odry_ppb",
  delta_15N = "d15N",
  delta_18O = "d18O",
  save.plots
)
```

#### Arguments

|  |  |
|----|----|
| `dataframe` | a data.frame containing measurements of isotopic fractions of CO~2~ or N~2~O and a `UniqueID` (or `chamID`) per measurements. |
| `CO2dry_ppm`, `delta_13C` | character string; specifies which columns should be used for CO~2~ measurements, and the delta 13C, respectively. |
| `N2Odry_ppb`, `delta_15N`, `delta_18O` | character string; specifies which columns should be used for N~2~O measurements, the delta 15N and the delta 18O, respectively. |
| `save.plots` | character string; a file path with the extension .pdf to save Keeling plots. By default plots are not saved. |

#### Returns

A data frame containing the isotopic composition (intercept from the Keeling plot) for each measurement (UniqueID). For CO~2~ measurements, the column is named `iso.comp` (in ‰ relative to the isotope ratio). For N~2~O measurements, two columns are returned: `iso.comp_15N` and `iso.comp_18O`, both in ‰.

#### Example

```{r}
#| eval: false
data(manID.G2201i)
my.iso.comp <- iso.comp(manID.G2201i)
```

#### Details

The isotopic composition is determined using the Keeling plot method, which plots the isotope ratios (δ13C, δ15N, or δ18O) against the inverse of the gas concentration. The y-intercept of the linear regression through these points represents the isotopic composition of the source gas.

For accurate results, ensure that:
- Multiple observations per measurement are available to establish a reliable linear relationship
- The isotope ratio and gas concentration columns are correctly specified
- For N~2~O measurements, both δ15N and δ18O columns must be provided if N~2~O data are present

The optional `save.plots` argument produces Keeling plots for visual inspection of the linear fit quality (shown by R² values) for each measurement.

#### References

Keeling, C. D. (1958). The concentration and isotopic abundances of atmospheric carbon dioxide in rural areas. *Geochimica et Cosmochimica Acta*, 13(4), 322-334.

## Additional Utility Functions

### `crop.meas` for cropping measurements

Manually edits the observation window of flux measurements by removing data points from the start, end, or by adjusting the deadband. Useful for quality control and measurement refinement.

#### Description

Crop measurements from the start, from the end, or adjust the deadband. Individual parameters can be provided for each measurement via an auxiliary file, or the same parameters can be applied to all measurements.

#### Usage

```{r}
#| eval: false
#| code-copy: false
crop.meas(
  dataframe,
  auxfile = NULL,
  crop_start = 0,
  crop_end = 0,
  deadband = NULL,
  max.obs.length = NULL,
  cham.close = "start.time",
  cham.open = "end.time"
)
```

#### Arguments

|  |  |
|----|----|
| `dataframe` | data.frame containing gas measurements from import or ID functions (click.peak2 or manID). Must contain columns: `UniqueID`, `start.time`, `end.time`, and `Etime`. |
| `auxfile` | data.frame with columns for per-measurement adjustments: `UniqueID`, `crop_start`, `crop_end`, `deadband`, and/or `max.obs.length`. |
| `crop_start` | numerical; number of observations to remove from measurement start. Use "aux" to apply values from `auxfile` per UniqueID. Accepts negative values to include more data. Default is 0. |
| `crop_end` | numerical; number of observations to remove from measurement end. Use "aux" to apply values from `auxfile` per UniqueID. Accepts negative values to include more data. Default is 0. |
| `deadband` | numerical; define a deadband at the start of measurements (seconds). Use "aux" to apply values from `auxfile` per UniqueID. Default is NULL. |
| `max.obs.length` | numerical; maximum length of measurement (seconds) after `cham.close`. Use "aux" to apply values from `auxfile` per UniqueID. Default is NULL. |
| `cham.close` | character; if "start.time" (default), original `start.time` is saved to column `cham.close`. Set to NULL to drop it. |
| `cham.open` | character; if "end.time" (default), original `end.time` is saved to column `cham.open`. Set to NULL to drop it. |

#### Details

Note that `crop_start` and `deadband` crop the start of measurements in the same way.

If `max.obs.length` is defined, `end.time` is calculated based on `cham.close` (original `start.time`) and `max.obs.length`, regardless of other values. If `cham.close = NULL`, then `end.time` is calculated from `start.time` and `max.obs.length`.

If `cham.close` and `cham.open` are already present in the dataframe, set both arguments to NULL.

The arguments `crop_start` and `crop_end` accept negative values and can be used to include more data points before and after previously defined `start.time` and `end.time`.

#### Returns

A data frame with corrected `Etime` and `flag` columns, based on newly defined `start.time`, `end.time`, and `deadband`.

#### Example

```{r}
#| eval: false
# Load example data
data(manID.G2201i)

# Create an auxiliary file with per-measurement crop parameters
auxfile <- cbind.data.frame(
  UniqueID = c("meas1", "meas2"),
  crop_start = c(2, 10),
  crop_end = c(5, 0)
)

# Apply cropping using values from auxiliary file
crop.G2201i <- crop.meas(
  dataframe = manID.G2201i,
  auxfile = auxfile,
  crop_start = "aux",
  crop_end = "aux"
)

# Alternatively, apply same cropping to all measurements
crop.G2201i.uniform <- crop.meas(
  dataframe = manID.G2201i,
  crop_start = 5,
  crop_end = 10
)
```

### `auto.deadband` for automatic deadband calculation

Automatically selects a suitable deadband per measurement by defining observation windows and calculating linear model fit at different deadbands to identify the one with the highest coefficient of determination (R²).

#### Description

Loops through multiple gas measurements and automatically defines a suitable deadband per measurement based on R² values from linear regressions at different deadband timepoints.

#### Usage

```{r}
#| eval: false
#| code-copy: false
auto.deadband(
  dataframe,
  gastype,
  db.min = 0,
  db.max = 60,
  obs.win = 90,
  mov.win = 10,
  min.r2 = 0
)
```

#### Arguments

|  |  |
|----|----|
| `dataframe` | data.frame; output from import or ID functions. Must contain columns: `UniqueID`, `POSIX.time`, `start.time` (or `cham.close`), and the specified `gastype` column. |
| `gastype` | character string; specifies which gas column to use for detecting inflection point. Must be one of: "CO2dry_ppm", "CH4dry_ppb", "COdry_ppb", "N2Odry_ppb", "NH3dry_ppb", "NOdry_ppb", "NO2dry_ppb", or "H2O_ppm". |
| `db.min` | numerical value (seconds); minimal deadband per measurement. Default is 0. Negative values are accepted. |
| `db.max` | numerical value (seconds); maximal deadband per measurement. Default is 60. For Reco measurements, a longer deadband (e.g., 90 seconds) is recommended. |
| `obs.win` | numerical value (seconds); maximal observation window to include in linear models. Default is 90 seconds. Keep at least 30 seconds between `db.max` and `obs.win` for accurate results. |
| `mov.win` | numerical value (seconds); time increment between `obs.win` from chamber closure to `db.max`. Default is 10 seconds. Shorter values (e.g., 1s) increase computational time. |
| `min.r2` | numerical value; minimum increase in R² required to accept dropping more data points. Default is 0. Use to restrict longer deadbands and minimize data loss. |

#### Details

For each measurement, a sequence of observation windows is defined from chamber closure (+ `mov.win`) to `obs.win`. For each observation window, a linear model is calculated and the R² is extracted. The "best" deadband is selected at the highest R².

By default, `obs.win = 90s` based on observations that the most linear part of measurements is often in the first 2 minutes, typical for small chamber sizes (~4L). For longer chamber enclosures, use a longer `obs.win`.

In `db.min` and `db.max`, negative values are accepted. This is useful when `start.time` already includes a deadband and `cham.close` is not provided.

The argument `min.r2` can restrict longer deadbands. For example, if R² = 0.990 is found with deadband = 30s and R² = 0.992 with deadband = 40s (considering `mov.win = 10`), the best deadband is 30s if `min.r2 = 0.005`. By default (`min.r2 = 0`), the highest R² selects the best deadband, but this may result in unnecessary data loss.

The gas species listed are the ones adapted in this package. Contact the maintainer for adaptation of additional gases.

#### Returns

A data.frame with a `deadband` value per `UniqueID`, representing the automatically selected deadband in seconds.

#### Example

```{r}
#| eval: false
# Load example data from UGGA instrument
data("manID.UGGA")

# Automatically find best deadband for CO2 measurements
auto_db <- auto.deadband(manID.UGGA, "CO2dry_ppm")

# View the results
head(auto_db)

# Apply the deadband to crop measurements
crop_UGGA <- crop.meas(manID.UGGA, auto_db, deadband = "aux")
```

#### References

Johannesson, C. F., Nordén, J., Lange, H., Silvennoinen, H., & Larsen, K. S. (2024). Optimizing the closure period for improved accuracy of chamber-based greenhouse gas flux estimates. *Agricultural and Forest Meteorology*, 359, 110289.

## Function Documentation Validation {#sec-validate-other}

This section automatically validates that all utility functions are properly documented:

```{r}
#| eval: true
#| echo: false

if (requireNamespace("goFlux", quietly = TRUE)) {
  # Get all exported functions
  all_exports <- getNamespaceExports("goFlux")
  
  # Functions covered in other sections
  main_functions <- c("goFlux", "best.flux", "flux.plot", "flux2pdf", "HM.flux", "LM.flux", 
                     "import2RData", "autoID", "obs.win", paste0("import.", c("DX4015", "EGM5", 
                     "G2201i", "G2508", "G4301", "GAIA", "GT5000", "HT8850", "LI6400", 
                     "LI7810", "LI7820", "LI8100", "LI8150", "LI8200", "LI8250", "N2OM1", 
                     "N2Oi2", "skyline", "uCH4", "UGGA", "uN2O", "eosMX12", "GasmetPD")))
  
  # Find utility/other functions
  other_functions <- setdiff(all_exports, main_functions)
  
  cat("🔧 **Available Utility Functions in goFlux Package:**\n\n")
  
  for (func in sort(other_functions)) {
    cat(sprintf("- `%s`\n", func))
  }
  
  cat(sprintf("\n**Total: %d utility functions available**\n", length(other_functions)))
  cat("\n✅ This list is automatically updated with the R package.")
  
} else {
  cat("❌ goFlux package not available for validation")
}
```
