---
title: "Other functions in the package goFlux"
format: 
  html: 
    css: styles.css
engine: knitr
editor: visual
---

Aside from the typical workflow of flux calculation, one may wish to estimate other parameters from gas measurements. Currently, this page only contains one additional function. To request more complementary functions, [open an issue on GitHub](https://github.com/Qepanna/goFlux/issues) or contact the maintainer of this package, Karelle Rheault ([karh\@ign.ku.dk](mailto:karh@ign.ku.dk){.email}), directly.

### `iso.comp` for isotopic composition determination

Extracts the y-intercept of the regression between the isotope ratios (13C, 15N or 18O) plotted againts the inverse of the gas concentration (CO~2~ or N~2~O).

#### Usage

```{r}
#| eval: false
#| code-copy: false
iso.comp(dataframe, CO2dry_ppm, delta_13C, N2Odry_ppb, delta_15N, delta_18O, save.plots)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `dataframe` | a data.frame containing measurements of isotopic fractions of \\ifelse{html}{\\out{CO<sub>2</sub>}}{\\eqn{CO\[2\]}{ASCII}} or \\ifelse{html}{\\out{N<sub>2</sub>O}}{\\eqn{N\[2\]O}{ASCII}} and a \\code{UniqueID} (or \\code{chamID}) per measurements. |
| `CO2dry\_ppm,delta\_13C` | character string; specifies which columns should be used for \\ifelse{html}{\\out{CO<sub>2</sub>}}{\\eqn{CO\[2\]}{ASCII}} measurements, and the delta 13C, respectively. |
| `N2Odry\_ppb,delta\_15N,delta\_18O` | character string; specifies which columns should be used for \\ifelse{html}{\\out{N<sub>2</sub>O}}{\\eqn{N\[2\]O}{ASCII}} measurements, the delta 15N and the delta 18O, respectively. |
| `save.plots` | character string; a file path with the extension .pdf to save Keeling plots. By default plots are not saved. |

#### Example

```{r}
#| eval: false

data(manID.G2201i)
my.iso.comp <- iso.comp(manID.G2201i)
```

### `auto.deadband` {#sec-auto-deadband}

#### Usage

```{r}
#| eval: false
#| code-copy: false
auto.deadband(dataframe, gastype, db.min, db.max, obs.win, mov.win, min.r2)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `dataframe` | data.frame; output from import or ID functions. |
| `gastype` | character string; specifies which gas column should be used to detect inflection point. Must be one of the following: "CO2dry\_ppm", "CH4dry\_ppb", "COdry\_ppb", "N2Odry\_ppb", "NH3dry\_ppb", "NOdry\_ppb", "NO2dry\_ppb" or "H2O\_ppm". |
| `db.min` | numerical value (seconds); minimal deadband per measurement. |
| `db.max` | numerical value (seconds); maximal deadband per measurement. |
| `obs.win` | numerical value (seconds); maximal observation window to include in the linear models. If possible, keept at least 30 seconds between \\code{db.max} and \\code{obs.win} for more accurate results. |
| `mov.win` | numerical value (seconds); time increment between \\code{obs.win}, from chamber closure to \\code{db.max}. |
| `min.r2` | numerical value; minimum increase in \\ifelse{html}{\\out{R<sup>2</sup>}}{\\eqn{R^{2}}{ASCII}} required to accept dropping more data points. |

#### Example

```{r}
#| eval: false

data("manID.UGGA")
auto_db <- auto.deadband(manID.UGGA, "CO2dry_ppm")
crop_UGGA <- crop.meas(manID.UGGA, auto_db, deadband = "aux")
```

### `autoID` {#sec-autoid}

#### Usage

```{r}
#| eval: false
#| code-copy: false
autoID(inputfile, auxfile, obs.length, deadband, crop.end, shoulder, drop)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `inputfile` | data.frame; output from import or align functions. |
| `auxfile` | data.frame; auxiliary data frame containing the columns \\code{start.time} and \\code{UniqueID}. \\code{start.time} must contain a date and be in POSIXct format. The time zone must be the same as the POSIX.time in \\code{inputfile}. The default time zone for the import functions is "UTC". A data frame from the Smart Chamber (LI-8200) can be used as an auxiliary file. In that case, \\code{chamID} will be used instead of \\code{UniqueID}, if \\code{UniqueID} cannot be found. |
| `obs.length` | numerical; chamber closure time (seconds). Default is NULL. If \\code{obs.length} is not provided, a column \\code{obs.length} should be contained in \\code{auxfile} or \\code{inputfile}. Alternatively, \\code{obs.length} will be calculated from \\code{start.time} and \\code{cham.open} or \\code{end.time} if found in \\code{auxfile} or \\code{inputfile}. |
| `deadband` | numerical; time (seconds) at the start of the measurement, after chamber closure, to be removed from the time stamps used to estimate the flux. |
| `crop.end` | numerical; time (seconds) at the end of the measurement, before chamber opening, to be removed from the time stamps used to estimate the flux. |
| `shoulder` | numerical; time before and after measurement in observation window (seconds). Default is 60 seconds. If the time between two measurements is smaller than \\code{shoulder}, the time stamps are duplicated. |
| `drop` | logical; if \\code{drop = TRUE} (default), time stamps without an assigned \\code{UniqueID} are removed from the final data frame. |

#### Example

```{r}
#| eval: false

# How to use in different situations:
library(dplyr)

## with a LGR instrument and an auxiliary file (.txt)
aux.path <- system.file("extdata", "aux_UGGA/aux_UGGA.txt", package = "goFlux")
auxfile <- read.delim(aux.path) %>%
  mutate(start.time = as.POSIXct(start.time, tz = "UTC"))
data(imp.UGGA)
autoID.UGGA <- autoID(inputfile = imp.UGGA, auxfile = auxfile, obs.length = 180)

## with a LI-COR instrument and the Smart Chamber as auxiliary file
data(imp.LI8200)
data(imp.LI7820)
autoID.LI7820 <- autoID(inputfile = imp.LI7820, deadband = 20, auxfile = imp.LI8200)
```

### `click.peak2` {#sec-click-peak2}

#### Usage

```{r}
#| eval: false
#| code-copy: false
click.peak2(ow.list, gastype, sleep, plot.lim, seq, warn.length, save.plots, width, height, abline, abline_corr)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `ow.list` | list of data.frame; output from the function \\code{\\link\[goFlux\]{obs.win}}. Must contain the columns \\code{gastype} (see below), \\code{POSIX.time} and \\code{UniqueID}. |
| `gastype` | character string; specifies which gas should be displayed on the plot to manually select start time and end time of measurements. Must be one of the following: "CO2dry\_ppm", "COdry\_ppb", "CH4dry\_ppb", "N2Odry\_ppb", "NO2dry\_ppb", "NOdry\_ppb", "NH3dry\_ppb" or "H2O\_ppm". Default is "CO2dry\_ppm". |
| `sleep` | numerical value; delay before closing the resulting plot. Grants a delay between measurements to visually inspect the output before processing the next measurement. Sleep must be shorter than 10 seconds. If \\code{sleep = NULL}, the plots will not close. |
| `plot.lim` | numerical vector of length 2; sets the Y axis limits in the plots. Default values are set for a typical gas measurement of "CO2dry\_ppm" from soils: \\code{plot.lim = c(380,1000)}. |
| `seq` | a numerical sequence that indicates objects in a list. By default, \\code{seq = NULL} and the function loops through all data frames in \\code{ow.list}. |
| `warn.length` | numerical value; limit under which a measurement is flagged for being too short (\\code{nb.obs < warn.length}). Default value is \\code{warn.length = 60}. |
| `save.plots` | character string; a file path with the extension .pdf to save the plots produced with click.peak2. By default, \\code{save.plot = NULL} and plots are not saved. |
| `width` | numerical value; width of the pop-up window. |
| `height` | numerical value; height of the pop-up window. |
| `abline` | logical; if TRUE (default), blue vertical lines indicate \\code{start.time} and \\code{end.time} on both plots. |
| `abline\_corr` | logical; if TRUE (default), red vertical lines indicate \\code{start.time\_corr} and \\code{end.time\_corr} on validation plot. |

#### Example

```{r}
#| eval: false

# IMPORTANT! This function makes use of the function graphics::identify()
# which is only supported on screen devices such as X11, windows and quartz.
# It is therefore essential to verify that your system options are compatible
# with this function before running it, to avoid errors. Here is an example
# of how to modify your system options for graphics device:
\dontrun{
default.device <- getOption("device") # save default option
options(device = "X11") # change system option to device = "X11"
options(device = default.device) # revert back to default option }

# There are multiple ways to use this function. Here are some examples:

library(dplyr)
\donttest{
## with a LGR instrument and an auxiliary file (.txt),
## and plots saved as pdf
aux.path <- system.file("extdata", "aux_UGGA/aux_UGGA.txt", package = "goFlux")
auxfile <- read.delim(aux.path) %>%
  mutate(start.time = as.POSIXct(start.time, tz = "UTC"))
data(imp.UGGA)
ow.UGGA <- obs.win(inputfile = imp.UGGA, auxfile = auxfile,
                   obs.length = 180, shoulder = 60)
save.to <- paste(tempdir(), "manID.UGGA.pdf", sep = "/")
manID.UGGA <- click.peak2(ow.UGGA, save.plots = save.to)
# In this example, plots are saved in a temporary directory using
# the function tempdir(). To save files in your working directory
# instead, use the function getwd().


## with a LI-COR instrument, the Smart Chamber as auxiliary file
## and only the first data frame selected
data(imp.LI8200)
data(imp.LI7820)
ow.LI7820 <- obs.win(inputfile = imp.LI7820, auxfile = imp.LI8200, shoulder = 60)
manID.LI7820 <- click.peak2(ow.LI7820, gastype = "N2Odry_ppb",
                            plot.lim = c(250,500), seq = 1)


## with the LI-6400, no auxiliary file
## and seq through the first 3 data frames
data(imp.LI6400)
ow.LI6400 <- obs.win(inputfile = imp.LI6400, shoulder = 0)
manID.LI6400 <- click.peak2(ow.LI6400, seq = seq(1,3)) }
```

### `crop.meas` {#sec-crop-meas}

#### Usage

```{r}
#| eval: false
#| code-copy: false
crop.meas(dataframe, auxfile, crop_start, crop_end, deadband, max.obs.length, cham.close, cham.open)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `dataframe` | a dataframe containing gas measurements, output from import or ID functions (click.peak2 or manID). Must contain the columns \\code{UniqueID}, \\code{start.time}, \\code{end.time} and \\code{Etime}. |
| `auxfile` | data.frame; auxiliary data frame containing the columns \\code{UniqueID}, \\code{crop\_start}, \\code{crop\_end}, \\code{deadband} and/or \\code{max.obs.length}. |
| `crop\_start` | numerical; number of observations to remove at the start of the measurement. To provide a unique value per UniqueID, the column \\code{crop\_start} must be present in the auxfile and the argument must be set to \\code{crop\_start = "aux"}. |
| `crop\_end` | numerical; number of observations to remove at the end of the measurement. To provide a unique value per UniqueID, the column \\code{crop\_end} must be present in the auxfile and the argument must be set to \\code{crop\_end = "aux"}. |
| `deadband` | numerical; define a deadband at the start of measurements (seconds). To provide a unique value per UniqueID, the column \\code{deadband} must be present in the auxfile and the argument must be set to \\code{deadband = "aux"}. |
| `max.obs.length` | numerical; maximum length of measurement (seconds) to keep after \\code{cham.close}. To provide a unique value per UniqueID, the column \\code{max.obs.length} must be present in the auxfile and the argument must be set to \\code{max.obs.length = "aux"}. |
| `cham.close` | character; if \\code{cham.close = "start.time"} (default), the original \\code{start.time} before cropping will be saved into a new column \\code{cham.close}. Set \\code{cham.close = NULL} to drop the original \\code{start.time}. |
| `cham.open` | character; if \\code{cham.open = "end.time"} (default), the original \\code{end.time} before cropping will be saved into a new column \\code{cham.open}. Set \\code{cham.open = NULL} to drop the original \\code{end.time}. |

#### Example

```{r}
#| eval: false

# Load example data
data(manID.G2201i)

# Create an auxfile to define different crop_start and crop_end per UniqueID
auxfile <- cbind.data.frame(UniqueID = c("meas1", "meas2"),
                            crop_start = c(2,10),
                            crop_end = c(5,0))

# The arguments "crop_start" and "crop_end" must be set to "aux" to use
# unique values from the auxfile
crop.G2201i <- crop.meas(dataframe = manID.G2201i, auxfile = auxfile,
                         crop_start = "aux", crop_end = "aux")

# Alternatively, cropping can be applied equally across all measurements
crop.G2201i <- crop.meas(dataframe = manID.G2201i, auxfile = auxfile,
                         crop_start = 5, crop_end = 5)
```

### `flux.term` {#sec-flux-term}

#### Usage

```{r}
#| eval: false
#| code-copy: false
flux.term(V_L, P_kPa, A_cm2, T_C, H2O_mol)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `V\_L` | numerical; total volume inside the chamber, tubes, instruments, etc. (L) |
| `P\_kPa` | numerical; atmospheric pressure (kPa) |
| `A\_cm2` | numerical; area of the soil surface inside the chamber \\ifelse{html}{\\out{(cm<sup>2</sup>)}}{\\eqn{(cm^2)}{ASCII}} |
| `T\_C` | numerical; air temperature before chamber closure (Celsius) |
| `H2O\_mol` | numerical; water vapor concentration in the air before chamber closure (mol/mol) |

#### Example

```{r}
#| eval: false
# No examples provided in roxygen.
```

### `g.factor` {#sec-g-factor}

#### Usage

```{r}
#| eval: false
#| code-copy: false
g.factor(HM.flux, LM.flux, eps)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `HM.flux` | a numerical value; flux estimate from a non-linear regression model (e.g. Hutchinson and Mosier). Calculated with the \\code{\\link\[goFlux\]{HM.flux}} function of this package. |
| `LM.flux` | a numerical value; flux estimate from a linear regression model. Calculated with the \\code{\\link\[goFlux\]{LM.flux}} function of this package. |

#### Example

```{r}
#| eval: false
# No examples provided in roxygen.
```

### `HM.flux` {#sec-hm-flux}

#### Usage

```{r}
#| eval: false
#| code-copy: false
HM.flux(gas.meas, time.meas, flux.term, k.max, k.min, Ct, C0, k.mult)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `gas.meas` | numerical vector containing gas measurements (ppm or ppb) |
| `time.meas` | numerical vector containing time stamps (seconds) |
| `flux.term` | numerical value; flux term calculated with the function \\code{\\link\[goFlux\]{flux.term}} |
| `Ct` | numerical value; gas concentration after \\emph{t} seconds (ppm or ppb). |
| `C0` | numerical value; initial gas concentration at time 0 second (ppm or ppb). |
| `k.max` | numerical value; kappa-max is the maximal curvature allowed in the Hutchinson and Mosier model. Calculated with the \\code{\\link\[goFlux\]{k.max}} function. |
| `k.min` | numerical value; a lower boundary value for kappa in the HM model. To allow for a log-like curvature, set \\code{k.min} below 0 (ex. -1), otherwise it should be just above 0 (ex. 10^-8). k.min cannot be 0 as this would result in a singular gradient. |
| `k.mult` | numerical value; a multiplier for the allowed k.max. Default is no multiplier (\\code{k.mult = 1}). |

#### Example

```{r}
#| eval: false
# No examples provided in roxygen.
```

### `k.max` {#sec-k-max}

#### Usage

```{r}
#| eval: false
#| code-copy: false
k.max(MDF, LM.flux, t)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `MDF` | numerical value; minimal detectable flux, calculated with the function \\code{\\link\[goFlux\]{MDF}} |
| `LM.flux` | numerical value; flux estimate from a linear flux calculation model, calculated with the function \\code{\\link\[goFlux\]{LM.flux}} |
| `t` | numerical value; measurement time (enclosure time) (seconds) |

#### Example

```{r}
#| eval: false
# No examples provided in roxygen.
```

### `LM.flux` {#sec-lm-flux}

#### Usage

```{r}
#| eval: false
#| code-copy: false
LM.flux(gas.meas, time.meas, flux.term)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `gas.meas` | numerical vector containing gas measurements (ppm or ppb) |
| `time.meas` | numerical vector containing time stamps (seconds) |
| `flux.term` | numerical value; flux term calculated with the function \\code{\\link\[goFlux\]{flux.term}} |

#### Example

```{r}
#| eval: false
# No examples provided in roxygen.
```

### `MDF` {#sec-mdf}

#### Usage

```{r}
#| eval: false
#| code-copy: false
MDF(p, t, flux.term)
```

#### Arguments

| Parameter | Description |
|:----------|:---------------------------|
| `p` | numerical; precision of the instrument (same units as measured gas; ex. ppm) |
| `t` | numerical; measurement time (enclosure time; seconds) |
| `flux.term` | numerical; flux term calculated with the function \\code{\\link\[goFlux\]{flux.term}} |

#### Example

```{r}
#| eval: false
# No examples provided in roxygen.
```

NA

