---
title: "Other functions in the package goFlux"
engine: knitr
editor: visual
---

Aside from the typical workflow of flux calculation, one may wish to estimate other parameters from gas measurements. Currently, this page only contains one additional function. To request more complementary functions, [open an issue on GitHub](https://github.com/Qepanna/goFlux/issues) or contact the maintainer of this package, Karelle Rheault ([karh\@ign.ku.dk](mailto:karh@ign.ku.dk){.email}), directly.

### `iso.comp` for isotopic composition determination

Extracts the y-intercept of the regression between the isotope ratios (13C, 15N or 18O) plotted againts the inverse of the gas concentration (CO~2~ or N~2~O).

#### Usage

::: callout-note
Code chunks under **Usage** sections are not part of the demonstration. They are meant to show you how to use the arguments in the function.
:::

```{r}
#| eval: false
#| code-copy: false
iso.comp(
  dataframe,
  CO2dry_ppm = "CO2dry_ppm",
  delta_13C = "Delta_Raw_iCO2",
  N2Odry_ppb = "N2Odry_ppb",
  delta_15N = "d15N",
  delta_18O = "d18O",
  save.plots
)
```

#### Arguments

|  |  |
|----|----|
| `dataframe` | a data.frame containing measurements of isotopic fractions of CO~2~ or N~2~O and a `UniqueID` (or `chamID`) per measurements. |
| `CO2dry_ppm`, `delta_13C` | character string; specifies which columns should be used for CO~2~ measurements, and the delta 13C, respectively. |
| `N2Odry_ppb`, `delta_15N`, `delta_18O` | character string; specifies which columns should be used for N~2~O measurements, the delta 15N and the delta 18O, respectively. |
| `save.plots` | character string; a file path with the extension .pdf to save Keeling plots. By default plots are not saved. |

#### Example

```{r}
#| eval: false
data(manID.G2201i)
my.iso.comp <- iso.comp(manID.G2201i)
```

## Additional Utility Functions

### `crop.meas` for cropping measurements

Crops data by selecting a time window within each UniqueID

#### Usage

```{r}
#| eval: false
#| code-copy: false
crop.meas(
  dataframe,
  start.cut,
  end.cut,
  UniqueID = "UniqueID"
)
```

#### Arguments

|  |  |
|----|----|
| `dataframe` | a data.frame containing gas measurements and a UniqueID per measurement. |
| `start.cut` | numerical value; time (in seconds) to start cropping measurements. |
| `end.cut` | numerical value; time (in seconds) to end cropping measurements. |
| `UniqueID` | character string; name of the column containing measurement IDs. |

#### Example

```{r}
#| eval: false
data(manID.UGGA)
cropped_data <- crop.meas(manID.UGGA, start.cut = 10, end.cut = 120)
```

### `auto.deadband` for automatic deadband calculation

Calculates an automatic deadband for gas measurements to remove initial non-steady-state data points

#### Usage

```{r}
#| eval: false
#| code-copy: false
auto.deadband(
  dataframe,
  gastype,
  UniqueID = "UniqueID",
  Etime = "Etime"
)
```

#### Arguments

|  |  |
|----|----|
| `dataframe` | a data.frame containing gas measurements and timestamps. |
| `gastype` | character string; column name of the gas measurements to analyze. |
| `UniqueID` | character string; name of the column containing measurement IDs. |
| `Etime` | character string; name of the column containing elapsed time. |

#### Example

```{r}
#| eval: false
data(manID.UGGA)
deadband_result <- auto.deadband(manID.UGGA, "CO2dry_ppm")
```

## Function Documentation Validation {#sec-validate-other}

This section automatically validates that all utility functions are properly documented:

```{r}
#| eval: true
#| echo: false

if (requireNamespace("goFlux", quietly = TRUE)) {
  # Get all exported functions
  all_exports <- getNamespaceExports("goFlux")
  
  # Functions covered in other sections
  main_functions <- c("goFlux", "best.flux", "flux.plot", "flux2pdf", "HM.flux", "LM.flux", 
                     "import2RData", "autoID", "obs.win", paste0("import.", c("DX4015", "EGM5", 
                     "G2201i", "G2508", "G4301", "GAIA", "GT5000", "HT8850", "LI6400", 
                     "LI7810", "LI7820", "LI8100", "LI8150", "LI8200", "LI8250", "N2OM1", 
                     "N2Oi2", "skyline", "uCH4", "UGGA", "uN2O", "eosMX12", "GasmetPD")))
  
  # Find utility/other functions
  other_functions <- setdiff(all_exports, main_functions)
  
  cat("🔧 **Available Utility Functions in goFlux Package:**\n\n")
  
  for (func in sort(other_functions)) {
    cat(sprintf("- `%s`\n", func))
  }
  
  cat(sprintf("\n**Total: %d utility functions available**\n", length(other_functions)))
  cat("\n✅ This list is automatically updated with the R package.")
  
} else {
  cat("❌ goFlux package not available for validation")
}
```
