---
title: "goAquaFlux"
subtitle: "Diffusive and ebullitive GHG flux partitioning for aquatic floating chambers"
engine: knitr
editor: visual
---

```{r}
#| include: false
source("scripts/autodoc.R")
```

## Overview

**goAquaFlux** extends the [goFlux](https://github.com/Qepanna/goFlux) R package
[@rheault2024] to aquatic floating-chamber measurements, where methane
reaches the atmosphere by two pathways with very different behaviour:

- **Diffusion** — a smooth, near-linear accumulation in the chamber headspace.
- **Ebullition** — bubbles, appearing as abrupt *steps* in concentration.

Standard chamber-flux software models a single smooth signal, so a naive fit
either discards bubble-affected incubations or silently folds ebullition into a
biased "diffusive" estimate. goAquaFlux detects the steps, fits the diffusive
flux on a bubble-free window using goFlux's validated LM/HM models and QA/QC,
and returns a partitioned total.

## How it works

The workflow is a pipeline of four steps, one function each, coordinated by
`goAquaFlux()` across all incubations in a dataset (split by `UniqueID`):

1. **Detect the bubbles.** `find.bubbles()` robustly standardises the
   concentration series (typically `CH4dry_ppb`) and scans it with a rolling
   variance, taken either over the first differences (`"diff"`, the default,
   robust to a steep diffusive trend) or over the concentration itself
   (`"variance"`, more sensitive to small steps on quiet traces). Contiguous
   periods above an adaptive threshold are flagged as bubbling events, and each
   event's size is then estimated as a local step on top of the diffusive trend.
2. **Fit the diffusive flux.** `goAquaFlux.diffusive()` fits the diffusive
   component on the bubble-free window (truncated at the first event) using
   goFlux's linear (LM) and non-linear (HM) models and the model-selection
   criteria. For non-bubble gases, the window is only cut at a bubble that
   coincides with an abrupt slope change.
3. **Compute the ebullitive flux.** `goAquaFlux.ebullition()` converts the
   summed bubble magnitudes into a flux,
   $F_E = \frac{\sum \Delta C_{\text{bubble}}}{t_{inc}} \times K$,
   with standard errors propagated from the per-event magnitudes, and returns
   an independent two-point endpoint estimate for diagnosis.
4. **Combine and check.** `goAquaFlux.total()` adds the two components,
   $F_T = F_E + F_D$, propagates the error assuming independence
   ($SE_T = \sqrt{SE_E^2 + SE_D^2}$), and flags results that are suspiciously
   larger than the endpoint estimate (e.g. a diffusive ramp mistaken for a
   bubble step).

`flux.plot.aqua()` then produces one diagnostic plot per incubation: the
concentration series, the shaded diffusive window, the detected bubbling
events, and a legend of the total / diffusive / ebullitive fluxes.

## Installation

```{r}
#| eval: false
# install.packages("remotes")
# The aquatic extension is part of goFlux from v0.3.2 onwards.
remotes::install_github("Qepanna/goFlux")
```

## Quick start

```{r}
#| eval: false
library(goFlux)

fluxres <- goAquaFlux(
  dataframe            = IDed,          # output of goFlux::autoID()
  gastype              = "CH4dry_ppb",
  use_bubble_detection = TRUE
)

# The detector is tunable. "diff" is the default; "variance" can be more
# sensitive to small bubbles on low-emission incubations.
fluxres <- goAquaFlux(
  dataframe            = IDed,
  gastype              = "CH4dry_ppb",
  use_bubble_detection = TRUE,
  bubble.method        = "variance",
  bubble.args          = list(min_magnitude = 10)
)

flux.plot.aqua(fluxres, IDed, gastype = "CH4dry_ppb",
               plot.display = c("diffusive.window", "ebullition.events"))
```

## Functions

### `goAquaFlux` for computing gas fluxes from aquatic chamber measurements {#goAquaFlux}

Main entry point: partitions aquatic GHG fluxes into diffusive and ebullitive
components across many incubations.

::: {.callout-note}
Code chunks under **Usage** sections are not part of the demonstration. They are meant to show how to use the function arguments.
:::

```{r}
#| echo: false
#| results: asis
cat(autodoc("goAquaFlux", level = 4))
```

### `find.bubbles` for detecting ebullition events {#find.bubbles}

Detects bubbling (ebullition) events as abrupt steps in a concentration
series.

```{r}
#| echo: false
#| results: asis
cat(autodoc("find.bubbles", level = 4))
```

### `goAquaFlux.diffusive` for estimating diffusive gas flux {#goAquaFlux.diffusive}

Diffusive flux on a bubble-free window, via goFlux's LM/HM models.

```{r}
#| echo: false
#| results: asis
cat(autodoc("goAquaFlux.diffusive", level = 4))
```

### `goAquaFlux.ebullition` for estimating ebullitive gas flux {#goAquaFlux.ebullition}

Ebullitive flux from detected bubble magnitudes.

```{r}
#| echo: false
#| results: asis
cat(autodoc("goAquaFlux.ebullition", level = 4))
```

### `goAquaFlux.total` for combining diffusive and ebullitive flux {#goAquaFlux.total}

Combine components with error propagation and consistency checks.

```{r}
#| echo: false
#| results: asis
cat(autodoc("goAquaFlux.total", level = 4))
```

### `flux.plot.aqua` for plotting aquatic chamber incubations {#flux.plot.aqua}

Per-incubation diagnostic plots of the diffusive window and ebullition events.

```{r}
#| echo: false
#| results: asis
cat(autodoc("flux.plot.aqua", level = 4))
```
