---
title: "Import raw data"
engine: knitr
editor: visual
format:
  html:
    toc: true
    toc-depth: 4
---
```{r}
#| include: false
source("scripts/autodoc.R")
```

The package contains a wide range of functions that let the user import raw data from a variety of instruments:

-   [**LI-COR trace gas analyzers**](https://www.licor.com/env/products/trace-gas/): LI-6400 for CO~2~ and H~2~O, LI-7810 for CO~2~, CH~4~ and H~2~O, LI-7820 for N~2~O and H~2~O, and LI-8100 for CO~2~ and H~2~O
-   [**LI-COR Smart Chamber (LI-8200)**](https://www.licor.com/env/products/soil-flux/survey): flexible survey system connected to any gas analyzer
-   [**LI-COR Multiplexer (LI-8250)**](https://www.licor.com/env/products/soil-flux/LI-8250): a single file that combine multiple gas analyzers and chambers
-   [**LI-COR Automated Flux System (LI-8100A + LI-8150)**](https://www.licor.com/products/soil-flux/LI-8100a): automated system for simple and reliable survey and long-term measurements of CO~2~
-   [**Los Gatos Research (ABB) laser gas analyzers**](https://new.abb.com/products/measurement-products/analytical/laser-gas-analyzers/laser-analyzers/lgr-icos-portable-analyzers): Ultra-portable Greenhouse Gas Analyzer (UGGA; GLA132 series), Microportable Greenhouse Gas Analyzer (MGGA; GLA131 series) for CO~2~, CH~4~ and H~2~O, GLA151-N2OM1 for N~2~O, CH~4~ and H~2~O, and N~2~O isotopic analyzer (GLA451-N2Oi2)
-   [**Picarro G2201*-i* Isotopic Analyzer**](https://www.picarro.com/environmental/products/g2201i_isotopic_analyzer): for δ^13^C for CH~4~ and CO~2~
-   [**Picarro G2508 Gas Concentration Analyzer**](https://www.picarro.com/g2508_gas_concentration_analyzer): for CO~2~, CH~4~, N~2~O, NH~3~ and H~2~O
-   [**Picarro GasScouter^TM^ G4301 Mobile Gas Concentration Analyzer**](https://www.picarro.com/products/gas-scouter-g4301): for CO~2~, CH~4~ and H~2~O
-   [**GAIA2TECH Automated ECOFlux chamber**](https://www.dmr.eu/technologies/gas-emission-measurements-eco2flux/automated-eco2flux-chamber/): for an easy import of data from any gas analyzer, and multiple chambers
-   [**Earth Bound Scientific SkyLine2D system**](https://www.earthboundscientific.com/products): raw files generated by GAIA2TECH
-   [**Gasmet DX4015 portable analyzer for humid conditions**](https://www.gasmet.com/products/category/portable-gas-analyzers/dx4015/): for CO, CO~2~, CH~4~, N~2~O, NH~3~ and H~2~O
-   [**Gasmet GT5000 Terra – Splashproof multigas FTIR analyzer**](https://www.gasmet.com/products/category/portable-gas-analyzers/gt5000-terra/): for CO, CO~2~, CH~4~, N~2~O, NH~3~ and H~2~O
-   [**PP-Systems EGM-5 Portable CO~2~ Gas Analyzer**](https://ppsystems.com/egm-5/): for CO~2~, O~2~ and H~2~O
-   [**Aeris Technologies MIRA Ultra Gas Analyzers**](https://aerissensors.com/ultra-series/): for N~2~O, CO~2~ and H~2~O or for CH~4~, C~2~H~6~ and H~2~O
-   [**Healthy Photon HT8850 all-in-one GHG analyzer**](https://en.healthyphoton.com/gongye/products/45.html): for CO~2~, CH~4~, N~2~O, and H~2~O
-   [**Eosense Portable Recirculating Multiplexer**](https://eosense.com/products/fluxlink.html): eosMX 12-Channel Autochamber Multiplexer
-   **Custom made multiplexer** from the University of Padova, Italy, which integrates data from multiple automatic chambers linked to a Gasmet DX4015.

There are two ways to import raw data files into R:

-   [Import all files](import.qmd#sec-import2RData) contained in a folder using the function `import2RData`.
-   [Import single files](import.qmd#sec-single) using a function unique to each instrument.

# import2RData {#sec-import2RData}

Import all raw gas measurement files contained in a folder (and all its sub-folders) using the function `import2RData`. Adapted to multiple greenhouse gas analyzers and other instruments:

-   [**LI-COR**](import.qmd#sec-import2RData-LICOR): LI-6400, LI-7810, LI-7820, LI-8100, LI-8100A + LI-8150 (automated system), LI-8200 (smart chamber), LI-8250 (Multiplexer)
-   [**Los Gatos Research (LGR)**](import.qmd#sec-import2RData-LGR): ultra-portable GGA (GLA132 series), micro ultra-portable GGA (GLA131 series), N2OM1 (GLA151 series) and N2Oi2 (GLA451 series)
-   [**GAIA2TECH (DMR)**](import.qmd#sec-import2RData-GAIA): automated chamber ECOFlux
-   [**EarthBound Scientigfic**](import.qmd#sec-import2RData-skyline): Skyline2D system
-   [**Gasmet**](import.qmd#sec-import2RData-Gasmet): DX4015 and GT5000
-   [**Picarro**](import.qmd#sec-import2RData-Picarro): G2201-*i*, G2508 and G4301
-   [**PP-Systems**](import.qmd#sec-import2RData-PP-System): EGM-5
-   [**Aeris Technologies**](import.qmd#sec-import2RData-Aeris): N~2~O/CO~2~ high accuracy analyzer and CH~4~/C~2~H~6~ natural gas detection system
-   [**Healthy Photon**](import.qmd#sec-import2RData-HealthyPhoton): HT8850
-   [**GasmetPD**](import.qmd#sec-import2RData-GasmetPD): Custom multiplexer from the University of Padova, Italy

::: {.callout-note}
Code chunks under **Usage** sections are not part of the demonstration. They are meant to show how to use the function arguments.
:::

```{r}
#| echo: false
#| results: asis
cat(autodoc("import2RData", level = 3, examples = FALSE))
```
## Examples

All files contained in the folder path specified are saved in a newly created folder "RData" in your working directory.

::: callout-important
In all examples, an example file from the package `goFlux` is retrieved using the function `system.file` to be used in the example. However, to use the function with your own data, the argument `path` must be the folder path containing your files:

```{r}
#| eval: false
import2RData(path = "your_raw_data_folder", instrument = "DX4015", 
             date.format = "ymd", prec = c(1.6, 23, 13, 2, 23, 33))
```
:::

::: callout-note
### Files in sub-folders

Note that all files with the right extension (i.e., .dat, .txt, etc.) will be converted to RData when there are sub-folders within the folder specified in the `path` argument of the function.

For example, the instrument G2201-i from Picarro normally saves raw files with a folder structure like such: year/month/day, with one (or sometimes more) file saved per daily folder. If one specifies in the function the folder path "month", all raw files from that month will be converted and saved in the RData folder created by the `import2RData` function, following the same folder structure as in the original folder path containing raw files.

See [the example from the instrument G2201-i from Picarro](#sec-import2RData-Picarro-G2201i) for a demonstration.
:::

::: callout-tip
### Load and merge files after import

If you wish to combine saved RData files after import, use the function `list.files` to retrieve all files within the RData folder with the pattern "imp.RData". Indeed, all files converted to RData with the import functions of the `goFlux` R package contain the pattern "imp" before the .RData extension. The argument `full.names = TRUE` allows for retrieval of files within sub-folders.

Then use the function `map_df` to get and load all files and combine them into one data frame:

```{r}
#| eval: false
all.imp <- list.files(path = "RData", pattern = "imp.RData", full.names = TRUE) %>%
  map_df(~ get(load(.x)))
```

Alternatively, use the argument `merge = TRUE` in the function `import2Rdata` to merge all saved files and assign them to an object directly.

Note that data frames can only be merged if they contain the same columns.
:::

### LI-COR {#sec-import2RData-LICOR}

#### LI-6400 {#sec-import2RData-LICOR-LI6400}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI6400", package = "goFlux")
import2RData(path = file.path, instrument = "LI-6400",
             date.format = "mdy", prec = c(0.15, 20))
```

#### LI-7810 {#sec-import2RData-LICOR-LI7810}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI7810", package = "goFlux")
import2RData(path = file.path, instrument = "LI-7810",
             date.format = "ymd", prec = c(3.5, 0.6, 45))
```

#### LI-7820 {#sec-import2RData-LICOR-LI7820}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI7820", package = "goFlux")
import2RData(path = file.path, instrument = "LI-7820",
             date.format = "ymd", prec = c(0.4, 45))
```

#### LI-8100 {#sec-import2RData-LICOR-LI8100}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI8100", package = "goFlux")
import2RData(path = file.path, instrument = "LI-8100",
             date.format = "ymd", prec = c(1, 10))
```

#### LI-8150 + LI-8100A automated system {#sec-import2RData-LICOR-LI8150}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI8150", package = "goFlux")
import2RData(path = file.path, instrument = "LI-8150",
             date.format = "ymd", prec = c(1, 10))
```

#### LI-8200 (Smart Chamber) {#sec-import2RData-LICOR-LI8200}

Note that with this instrument, `keep_all` and `prec` are not valid arguments.

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI8200", package = "goFlux")
import2RData(path = file.path, instrument = "LI-8200", date.format = "ymd")

```

#### LI-8250 (Multiplexer) {#sec-import2RData-LICOR-LI8250}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/LI8250", package = "goFlux")
import2RData(path = file.path, instrument = "LI-8250", date.format = "ymd",
             inst1 = "LI-7810", gas1 = c("CO2_DRY", "CH4_DRY", "H2O"),
             prec1 = c(3.5, 0.6, 45), SWC.col = "SWC_1", Tsoil.col = "TS_1")
```

### LGR {#sec-import2RData-LGR}

#### UGGA (GLA132 series) {#sec-import2RData-LGR-UGGA}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/UGGA", package = "goFlux")
import2RData(path = file.path, instrument = "UGGA",
             date.format = "dmy", prec = c(0.2, 1.4, 50))
```

#### MGGA (GLA131 series) {#sec-import2RData-LGR-MGGA}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/UGGA", package = "goFlux")
import2RData(path = file.path, instrument = "UGGA",
             date.format = "dmy", prec = c(0.35, 0.9, 200))
```

#### N2OM1 (GLA151 series) {#sec-import2RData-LGR-N2OM1}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/N2OM1", package = "goFlux")
import2RData(path = file.path, instrument = "N2OM1",
             date.format = "dmy", prec = c(0.5, 2, 50))
```

#### N2Oi2 (GLA451 series) {#sec-import2RData-LGR-N2Oi2}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/N2Oi2", package = "goFlux")
import2RData(path = file.path, instrument = "N2Oi2",
             date.format = "dmy", prec = c(0.05, 50))
```

### GAIA2TECH {#sec-import2RData-GAIA}

#### Automated chamber ECOFlux

Note that with this instrument, `keep_all` is not a valid argument. However, several other arguments can be used: `active`, `pivot`, `flag`, `background`, etc. Look up the [GAIA2TECH ECOFlux](#sec-single-GAIA) instrument under the "Single file import" section for more details.

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/GAIA", package = "goFlux")
import2RData(path = file.path, instrument = "GAIA", date.format = "ymd",
             pivot = "long", active = TRUE, flag = c(7,11),
             background = FALSE,
             CH.col = "COM5A0",
             SWC.col = "1C08_Soil Moisture",
             Tsoil.col = "1C07_Soil Temperature",
             Tcham.col = "2C07_Chamber Temperature",
             PAR.col = "3C07_Sunlight",
             WTD.col = NULL,
             Op.stat.col = "0C06_OperatingStatus",
             inst1 = "XT2C00_Instrument",
             inst2 = "XT3C00_Instrument",
             gas1 = c("XT2C04_CH4", "XT2C05_CO2", "XT2C06_H2O"),
             gas2 = c("XT3C04_N2O", "XT3C05_H2O"),
             prec1 = c(0.6, 3.5, 45),
             prec2 = c(0.4, 45))
```

#### Skyline2D from Earth Bound Scientific {#sec-import2RData-skyline}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/skyline", package = "goFlux")
import2RData(path = file.path, instrument = "skyline", date.format = "ymd",
             background = FALSE,
             CH.col = "CH ID",
             CH.clo.col = "Chamber closed",
             sensor1 = "Analog Sensor1",
             sensor2 = "Analog Sensor2",
             inst = "Instrument",
             gas = c("CO2", "CH4 Dry", "N2O Dry", "H2O"),
             prec = c(0.24, 0.3, 5, 500))
```

### Gasmet {#sec-import2RData-Gasmet}

#### DX4015 {#sec-import2RData-Gasmet-DX4015}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/DX4015", package = "goFlux")
import2RData(path = file.path, instrument = "DX4015",
             date.format = "ymd", prec = c(1.6, 23, 13, 2, 23, 33))
```

#### GT5000 {#sec-import2RData-Gasmet-GT5000}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/GT5000", package = "goFlux")
import2RData(path = file.path, instrument = "GT5000",
             date.format = "ymd", prec = c(1.6, 23, 13, 2, 23, 33))
```

### Picarro {#sec-import2RData-Picarro}

#### G2201-i {#sec-import2RData-Picarro-G2201i}

The arguments `CH4`, `CO2`, `sum` and `range` are used with the function `import.G2201i` only. Look up the [Picarro G2201-i](#sec-single-Picarro-G2201i) instrument under the "Single file import" section for more details.

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/G2201i", package = "goFlux")
import2RData(path = file.path, instrument = "G2201i",
             date.format = "ymd", prec = c(0.2, 50, 100))
```

#### G2508 {#sec-import2RData-Picarro-G2508}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/G2508", package = "goFlux")
import2RData(path = file.path, instrument = "G2508",
             date.format = "ymd", prec = c(0.24, 0.3, 5, 0.16, 500))
```

#### G4301 {#sec-import2RData-Picarro-G4301}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/G4301", package = "goFlux")
import2RData(path = file.path, instrument = "G4301",
             date.format = "ymd", prec = c(0.025, 0.1, 10))
```

### PP-Systems {#sec-import2RData-PP-System}

#### EGM-5 {#sec-import2RData-PP-System-EGM5}

The argument `proc.data.field` is used with the PP-Sytems EGM-5 instrument only. Look up the [EGM-5 Operation Manual](https://ppsystems.com/download/technical_manuals/80109-1-EGM-5_Operation_V103.pdf) at page 90 for more details about the different Process Data Fields (`proc.data.field`), or the [PP-Systems EGM-5](#sec-single-PP-System) instrument under the "Single file import" section for more details about other arguments used with this instrument.

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/EGM5", package = "goFlux")
# The argument proc.data.field is used with the PP-System EGM-5 instrument only.
import2RData(path = file.path, instrument = "EGM5",
             date.format = "dmy", prec = c(3, 1, 500), proc.data.field = 2)
```

### Aeris Technologies {#sec-import2RData-Aeris}

#### MIRA Ultra CH~4~/ C~2~H~6~ analyzer {#sec-import2RData-Aeris-uCH4}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/uCH4", package = "goFlux")
import2RData(path = file.path, instrument = "uCH4",
             date.format = "mdy", prec = c(1, 0.5, 15))
```

#### MIRA Ultra N~2~O/ CO~2~ analyzer {#sec-import2RData-Aeris-uN2O}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/uN2O", package = "goFlux")
import2RData(path = file.path, instrument = "uN2O",
             date.format = "mdy", prec = c(0.2, 0.2, 15))
```

### Healthy Photon {#sec-import2RData-HealthyPhoton}

#### HT8850 all-in-one GHG analyzer {#sec-import2RData-HealthyPhoton-HT8850}

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/HT8850", package = "goFlux")
import2RData(path = file.path, instrument = "HT8850",
             date.format = "ymd", prec = c(0.5, 3, 0.5, 10))
```

### GasmetPD {#sec-import2RData-GasmetPD}

#### Custom made multiplexer linked with the Gasmet DX4015

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata/GasmetPD", package = "goFlux")
import2RData(path = file.path, instrument = "GasmetPD",
             gas = c("CO2_ppm", "CH4_ppm", "N2O_ppm", "NH3_ppm", "H2O_pct"),
             date.format = "ymd", prec = c(1.6, 13, 2, 23, 33))
```

# Single file import {#sec-single}

Each instrument has a unique function for single file import. The advantage of using this function is that a file can be saved in the environment in an object instead of being saved as an .RData file in the working directory. Just like the `import2RData` function, there is a function adapted to multiple greenhouse gas analyzers and other instruments:

-   [**LI-COR**](import.qmd#sec-single-LICOR): LI-6400, LI-7810, LI-7820, LI-8100, LI-8100A + LI-8150 (automated system), LI-8200 (smart chamber), LI-8250 (Multiplexer)
-   [**Los Gatos Research (LGR)**](import.qmd#sec-single-LGR): ultra-portable GGA (GLA132 series), micro ultra-portable GGA (GLA131 series), N2OM1 (GLA151 series) and N2Oi2 (GLA451 series)
-   [**GAIA2TECH (DMR)**](import.qmd#sec-single-GAIA): automated chamber ECOFlux
-   [**EarthBound Scientific**](import.qmd#sec-single-skyline): Skyline2D system
-   [**Gasmet**](import.qmd#sec-single-Gasmet): DX4015 and GT5000
-   [**Picarro**](import.qmd#sec-single-Picarro): G2201-*i*, G2508 and G4301
-   [**PP-Systems**](import.qmd#sec-single-PP-System): EGM-5
-   [**Aeris Technologies**](import.qmd#sec-single-Aeris): N~2~O/CO~2~ high accuracy analyzer and CH~4~/C~2~H~6~ natural gas detection system
-   [**Healthy Photon**](import.qmd#sec-single-HealthyPhoton): HT8850
-   [**GasmetPD**](import.qmd#sec-single-GasmetPD): Custom multiplexer from the University of Padova, Italy

## LI-COR {#sec-single-LICOR}

### LI-6400 {#sec-single-LICOR-LI6400}

Import single raw gas measurement files from the LI-COR 6400 (CO~2~ and H~2~O GHG analyzer) with the function `import.LI6400`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI6400", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI6400/LI6400.txt", package = "goFlux")

# Import in the environment
imp.LI6400 <- import.LI6400(inputfile = file.path)

# Or save in a RData folder in the working directory
import.LI6400(inputfile = file.path, save = TRUE)
```

### LI-7810 {#sec-single-LICOR-LI7810}

Import single raw gas measurement files from the LI-COR 7810 (CO~2~, CH~4~ and H~2~O GHG analyzer) with the function `import.LI7810`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI7810", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI7810/LI7810.data", package = "goFlux")

# Import in the environment
imp.LI7810 <- import.LI7810(inputfile = file.path)

# Or save in a RData folder in the working directory
import.LI7810(inputfile = file.path, save = TRUE)
```

### LI-7820 {#sec-single-LICOR-LI7820}

Import single raw gas measurement files from the LI-COR 7820 (N~2~O and H~2~O GHG analyzer) with the function `import.LI7820`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI7820", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI7820/LI7820.data", package = "goFlux")

# Import in the environment
imp.LI7820 <- import.LI7820(inputfile = file.path)

# Or save in a RData folder in the working directory
import.LI7820(inputfile = file.path, save = TRUE)
```

### LI-8100 {#sec-single-LICOR-LI8100}

Import single raw gas measurement files from the LI-COR 8100 (CO~2~ and H~2~O GHG analyzer) with the function `import.LI8100`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI8100", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI8100/LI8100.81x", package = "goFlux")

# Import in the environment
imp.LI8100 <- import.LI8100(inputfile = file.path)

# Or save in a RData folder in the working directory
import.LI8100(inputfile = file.path, save = TRUE)
```

### LI-8150 + LI-8100A {#sec-single-LICOR-LI8150}

Import single raw gas measurement files from the LI-COR Multiplexer LI-8150 connected with up to 16 chambers and the LI-8100A gas analyzer (CO~2~ and H~2~O) with the function `import.LI8150`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI8150", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI8150/LI8150.81x", package = "goFlux")

# Import in the environment
imp.LI8150 <- import.LI8150(inputfile = file.path)

# Or save in a RData folder in the working directory
import.LI8150(inputfile = file.path, save = TRUE)
```

### LI-8200 (Smart Chamber) {#sec-single-LICOR-LI8200}

Import single raw data files from the LI-COR smart chamber (LI-8200) with the function `import.LI8200`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI8200", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI8200/LI8200.json", package = "goFlux")

# Import in the environment
imp.LI8200 <- import.LI8200(inputfile = file.path)

# Or save in a RData folder in the working directory
import.LI8200(inputfile = file.path, save = TRUE)
```

### LI-8250 (Multiplexer) {#sec-single-LI8250}

Import single raw data files from the LI-COR Multiplexer (LI-8250) with the function `import.LI8250`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.LI8250", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "LI8250/LI8250.82z", package = "goFlux")

# Import in the environment
imp.LI8250 <- import.LI8250(inputfile = file.path, inst1 = "LI-7810",
                            gas1 = c("CO2_DRY", "CH4_DRY", "H2O"),
                            prec1 = c(3.5, 0.6, 45), SWC.col = "SWC_1",
                            Tsoil.col = "TS_1")

# Or save in a RData folder in the working directory
import.LI8250(inputfile = file.path, save = TRUE, inst1 = "LI-7810",
              gas1 = c("CO2_DRY", "CH4_DRY", "H2O"),
              prec1 = c(3.5, 0.6, 45), SWC.col = "SWC_1",
              Tsoil.col = "TS_1")
```

## LGR {#sec-single-LGR}

### UGGA and MGGA {#sec-single-LGR-UGGA}

Import single raw gas measurement files from the ultra-portable GHG analyzers (GLA132-UGGA and GLA131-MGGA) from Los Gatos Research (CO~2~, CH~4~ and H~2~O) with the function `import.UGGA`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.UGGA", level = 4, examples = FALSE))
```
#### Examples

##### UGGA (GLA132 series)

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "UGGA/UGGA.txt", package = "goFlux")

# Import in the environment
imp.UGGA <- import.UGGA(inputfile = file.path, prec = c(0.2, 1.4, 50))

# Or save in a RData folder in the working directory
import.UGGA(inputfile = file.path, prec = c(0.2, 1.4, 50), save = TRUE)
```

##### MGGA (GLA131 series)

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "UGGA/UGGA.txt", package = "goFlux")

# Import in the environment
imp.MGGA <- import.UGGA(inputfile = file.path, prec = c(0.35, 0.9, 200))

# Or save in a RData folder in the working directory
import.UGGA(inputfile = file.path, prec = c(0.35, 0.9, 200), save = TRUE)
```

### N2OM1 {#sec-single-LGR-N2OM1}

Import single raw gas measurement files from the N2OM1 GHG analyzers (GLA151 series) from Los Gatos Research (N~2~O, CH~4~ and H~2~O) with the function `import.N2OM1`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.N2OM1", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "N2OM1/N2OM1.txt", package = "goFlux")

# Import in the environment
imp.N2OM1 <- import.N2OM1(inputfile = file.path, prec = c(0.5, 2, 50))

# Or save in a RData folder in the working directory
import.N2OM1(inputfile = file.path, prec = c(0.5, 2, 50), save = TRUE)
```

### N2Oi2 {#sec-single-LGR-N2Oi2}

Import single raw gas measurement files from the Isotopic N~2~O analyzer (GLA451 series) from Los Gatos Research (N~2~O and site-specific isotopic ratios of N and O) with the function `import.N2Oi2`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.N2Oi2", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "N2Oi2/N2Oi2.txt", package = "goFlux")

# Import in the environment
imp.N2Oi2 <- import.N2Oi2(inputfile = file.path, prec = c(0.05, 50))

# Or save in a RData folder in the working directory
import.N2Oi2(inputfile = file.path, prec = c(0.05, 50), save = TRUE)
```

## GAIA2TECH {#sec-single-GAIA}

### Automated chamber ECOFlux

With the function `import.GAIA`, import single raw gas measurement files from the automated chamber ECOFlux (GAIA2TECH) linked with up to three gas analyzers. Default settings are set for two LI-COR instruments: LI-7810 (CO~2~, CH~4~ and H~2~O) and LI7820 (N~2~O and H~2~O), to match the example data file provided with the package.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.GAIA", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "GAIA/GAIA.csv", package = "goFlux")

# Import in the environment
imp.GAIA <- import.GAIA(inputfile = file.path)

# Or save in a RData folder in the working directory
import.GAIA(inputfile = file.path, save = TRUE)
```

### SkyLine2D from Earth Bound Scientific {#sec-single-skyline}

Imports single raw gas measurement files from the SkyLine2D system from Earth Bound Scientific from the datalog file created by GAIA2TECH. Default settings are set for the Picarro G2508 (CO~2~, CH~4~, N~2~O and H~2~O), to match the example data file provided with the package.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.skyline", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "skyline/skyline.csv", package = "goFlux")

# Import in the environment
imp.skyline <- import.skyline(inputfile = file.path)

# Or save in a RData folder in the working directory
import.skyline(inputfile = file.path, save = TRUE)
```

## Gasmet {#sec-single-Gasmet}

### DX4015 {#sec-single-Gasmet-DX4015}

Import single raw gas measurement files from the Gasmet DX4015 gas analyzer (CO, CO~2~, CH~4~, N~2~O, NH~3~ and H~2~O) with the function `import.DX4015`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.DX4015", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "DX4015/DX4015.TXT", package = "goFlux")

# Import in the environment
imp.DX4015 <- import.DX4015(inputfile = file.path)

# Or save in a RData folder in the working directory
import.DX4015(inputfile = file.path, save = TRUE)
```

### GT5000 {#sec-single-Gasmet-GT5000}

Import single raw gas measurement files from the Gasmet GT5000 gas analyzer (CO, CO~2~, CH~4~, N~2~O, NH~3~ and H~2~O) with the function `import.GT5000`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.GT5000", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "GT5000/GT5000.TXT", package = "goFlux")

# Import in the environment
imp.GT5000 <- import.GT5000(inputfile = file.path)

# Or save in a RData folder in the working directory
import.GT5000(inputfile = file.path, save = TRUE)
```

## Picarro {#sec-single-Picarro}

### G2201-i {#sec-single-Picarro-G2201i}

Import single raw gas measurement files from the Picarro G2201-*i* (CO~2~ and CH~4~ isotopic analyzer) with the function `import.G2201i`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.G2201i", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "G2201i/2025/05/13/G2201i.dat", package = "goFlux")

# Import in the environment
imp.G2201i <- import.G2201i(inputfile = file.path)

# Or save in a RData folder in the working directory
import.G2201i(inputfile = file.path, save = TRUE)
```

### G2508 {#sec-single-Picarro-G2508}

Import single raw gas measurement files from the Picarro G2508 (CO~2~, CH~4~, N~2~O, NH~3~, and H~2~O GHG analyzer) with the function `import.G2508`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.G2508", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "G2508/2022/08/01/G2508.dat", package = "goFlux")

# Import in the environment
imp.G2508 <- import.G2508(inputfile = file.path)

# Or save in a RData folder in the working directory
import.G2508(inputfile = file.path, save = TRUE)
```

### G4301 {#sec-single-Picarro-G4301}

Import single raw gas measurement files from the Picarro G4301 (CO~2~, CH~4~ and H~2~O GHG analyzer) with the function `import.G4301`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.G4301", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "G4301/G4301.dat", package = "goFlux")

# Import in the environment
imp.G4301 <- import.G4301(inputfile = file.path)

# Or save in a RData folder in the working directory
import.G4301(inputfile = file.path, save = TRUE)
```

## PP-Systems {#sec-single-PP-System}

### EGM-5 {#sec-single-PP-System-EGM5}

Import single raw gas measurement files from the PP-Systems EGM-5 Portable CO~2~ Gas Analyzer, with the function `import.EGM5`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.EGM5", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "EGM5/EGM5.TXT", package = "goFlux")

# Import in the environment
imp.EGM5 <- import.EGM5(inputfile = file.path)

# Or save in a RData folder in the working directory
import.EGM5(inputfile = file.path, save = TRUE)
```

## Aeris Technologies {#sec-single-Aeris}

### MIRA Ultra CH~4~/ C~2~H~6~ analyzer {#sec-single-Aeris-uCH4}

Import single raw data files from the MIRA Ultra CH~4~/ C~2~H~6~ Natural Gas Leak Detection System (with GPS) from Aeris Technologies with the function `import.uCH4`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.uCH4", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "uCH4/uCH4.txt", package = "goFlux")

# Import in the environment
imp.uCH4 <- import.uCH4(inputfile = file.path)

# Or save in a RData folder in the working directory
import.uCH4(inputfile = file.path, save = TRUE)
```

### MIRA Ultra N~2~O/ CO~2~ analyzer {#sec-single-Aeris-uN2O}

Import single raw data files from the MIRA Ultra N~2~O/ CO~2~ High Accuracy Analyzer (with GPS) from Aeris Technologies with the function `import.uN2O`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.uN2O", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "uN2O/uN2O.txt", package = "goFlux")

# Import in the environment
imp.uN2O <- import.uN2O(inputfile = file.path)

# Or save in a RData folder in the working directory
import.uN2O(inputfile = file.path, save = TRUE)
```

## Healthy Photon {#sec-single-HealthyPhoton}

### HT8850 all-in-one GHG analyzer {#sec-single-HealthyPhoton-HT8850}

Imports single raw gas measurement files from the all-in-one GHG analyzer HT8850 from Healthy Photon (CO~2~, CH~4~, N~2~O and H~2~O)

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.HT8850", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "HT8850/HT8850.txt", package = "goFlux")

# Import in the environment
imp.HT8850 <- import.HT8850(inputfile = file.path)

# Or save in a RData folder in the working directory
import.HT8850(inputfile = file.path, save = TRUE)
```

## GasmetPD {#sec-single-GasmetPD}

### Custom made multiplexer linked with the Gasmet DX4015

Imports single raw gas measurement files from the GasmetPD, a custom made multiplexer linked with the Gasmet DX4015 (CO~2~, CH~4~, N~2~O, NO~2~, NO, NH~3~ and H~2~O)

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.GasmetPD", level = 4, examples = FALSE))
```
#### Examples

```{r}
#| eval: false
# Retrieve file path from example file in the goFlux package
# using the function system.file
file.path <- system.file("extdata", "GasmetPD/GasmetPD.txt", package = "goFlux")

# Import in the environment
imp.GasmetPD <- import.GasmetPD(inputfile = file.path, prec = c(1.6, 13, 2, 23, 33))

# Or save in a RData folder in the working directory
import.GasmetPD(inputfile = file.path, prec = c(1.6, 13, 2, 23, 33), save = TRUE)
```

## Eosense {#sec-single-Eosense}

### eosMX12 12-Channel Multiplexer {#sec-single-eosMX12}

Import single raw data files from the Eosense eosMX12 12-channel multiplexer with the function `import.eosMX12`.

```{r}
#| echo: false
#| results: asis
cat(autodoc("import.eosMX12", level = 4))
```