---
title: "goAquaFlux"
subtitle: "Diffusive and ebullitive GHG flux partitioning for aquatic floating chambers"
toc: false
---

## Overview

**aquaGHG** extends the [goFlux](https://github.com/Qepanna/goFlux) R package
(Rheault et al., 2024) to aquatic floating-chamber measurements, where methane
reaches the atmosphere by two pathways with very different behaviour:

- **Diffusion** — a smooth, near-linear accumulation in the chamber headspace.
- **Ebullition** — bubbles, appearing as abrupt *steps* in concentration.

Standard chamber-flux software models a single smooth signal, so a naive fit
either discards bubble-affected incubations or silently folds ebullition into a
biased "diffusive" estimate. aquaGHG detects the steps, fits the diffusive flux
on a bubble-free window using goFlux's validated LM/HM models and QA/QC, and
returns a partitioned total.

## Installation

```{r}
#| eval: false
# install.packages("remotes")
remotes::install_github("Qepanna/goFlux")          # the parent package
remotes::install_github("Qepanna/goFlux@aquaghg")    # this fork + aqua extension
```

## Quick start

```{r}
#| eval: false
library(goFlux)

fluxres <- goAquaFlux(
  dataframe            = IDed,          # output of goFlux::autoID()
  gastype              = "CH4dry_ppb",
  use_bubble_detection = TRUE
)

flux.plot.aqua(fluxres, IDed, gastype = "CH4dry_ppb",
               plot.display = c("diffusive.window", "ebullition.events"))
```

## Where to go next

- [Getting started](articles/getting-started.qmd) — a full worked example.
- [Bubble detection](articles/bubble-detection.qmd) — how `find.bubbles()` works and how to tune it.
- [The partitioning workflow](articles/partitioning-workflow.qmd) — what happens between detection and the final flux.
- [Reference](reference/index.qmd) — every function, argument by argument.

## Citation

If you use aquaGHG, please cite both the parent package and the aquatic
extension:

> Rheault, K., Christiansen, J. R., & Larsen, K. S. (2024). goFlux: A
> user-friendly way to calculate GHG fluxes yourself, regardless of user
> experience. *Journal of Open Source Software*, 9(96), 6393.

> *[Add your aquaGHG reference once the MEE paper is submitted.]*
