---
title: "Other functions in the package goFlux"
format: 
  html: 
    css: styles.css
engine: knitr
editor: visual
---
```{r}
#| include: false
source("scripts/autodoc.R")
```

Aside from the typical workflow of flux calculation, one may wish to estimate other parameters from gas measurements. Currently, this page only contains one additional function. To request more complementary functions, [open an issue on GitHub](https://github.com/Qepanna/goFlux/issues) or contact the maintainer of this package, Karelle Rheault ([karh\@ign.ku.dk](mailto:karh@ign.ku.dk){.email}), directly.

### `iso.comp` for isotopic composition determination

Extracts the y-intercept of the regression between the isotope ratios (13C, 15N or 18O) plotted againts the inverse of the gas concentration (CO~2~ or N~2~O).

```{r}
#| echo: false
#| results: asis
cat(autodoc("iso.comp", level = 4))
```

#### Examples

```{r}
#| eval: false
data(manID.G2201i)
my.iso.comp <- iso.comp(manID.G2201i)
```
