---
title: "Other functions in the package goFlux"
format: 
  html: 
    css: styles.css
engine: knitr
editor: visual
---
```{r}
#| include: false
source("scripts/autodoc.R")
```

Aside from the typical workflow of flux calculation, one may wish to estimate other parameters from gas measurements. Currently, this page only contains a handful functions. To request more complementary functions, [open an issue on GitHub](https://github.com/Qepanna/goFlux/issues) or contact the maintainer of this package, Karelle Rheault ([karh\@ign.ku.dk](mailto:karh@ign.ku.dk){.email}), directly.

### `crop.meas` for additional modifications of the observation window {#crop.meas}

Crop measurements from the start, from the end, or adjust the deadband. Individual parameters can be provided for each measurement, in an auxiliary file, or the same parameters can be applied to all measurements.

::: {.callout-note}
Code chunks under **Usage** sections are not part of the demonstration. They are meant to show how to use the function arguments.
:::

```{r}
#| echo: false
#| results: asis
cat(autodoc("crop.meas", level = 4))
```

### `iso.comp` for isotopic composition determination

Extracts the y-intercept of the regression between the isotope ratios (13C, 15N or 18O) plotted againts the inverse of the gas concentration (CO~2~ or N~2~O).

```{r}
#| echo: false
#| results: asis
cat(autodoc("iso.comp", level = 4))
```

### `auto.deadband` for an automatic selection of deadband per measurement {#auto.deadband}
```{r}
#| echo: false
#| results: asis
cat(autodoc("auto.deadband", level = 4))
```
